Waving and swaying

### 2014-08-11 Paper/grass.tex
More changes to pages 1 and 2

\begin{document}
\title{Waving and swaying}
\author{Ravi Singh, Shreyas Mandre}
\affiliation{Brown University, Providence RI 02912 USA}
\author{L. Mahadevan}
\affiliation{Harvard University, Cambridge MA 02138 USA}
\author{Mahesh Bandi\footnote{work performed while visiting Brown University}}
\affiliation{Okinawa Institute of Science and Technology, Okinawa, Japan}
\author{Amala Mahadevan}
\affiliation{Woods Hold Institute of Oceanography, Woods Hole MA USA}
%\author{Shreyas Mandre}
%\affiliation{Brown University, Providence RI 02912 USA}

@@

\section{Introduction}
Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as 
transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc \cite{Fonseca87,Nepf99}. 
The response of flexible grass canopies to steady current in form of large amplitude coherent oscillation, known as mo-nami \cite{AckermanOkubo93}, plays a central role
in the recruitment of microscopic marine organism such as blue mussel larvae \cite{Grizzle96}. The hydrodynamic mechanism underlying monami is the focus of this paper. 
Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically. Another major
difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies.
Due to these differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domian is applicable only to marine canopies. 
Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to
different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to 
Kelvin-Helmholtz instability. Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
%Sea grasses exhibit large amplitude wave like motion in response to these eddies propagating downstream.
%are exhibited in form of large amplitude oscillations knows as mo-nami.
While shear layer model successfully predict frequency of mo-nami for number of experimental observations,
several aspect of existing theory remain unexplained. First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self 
consistent theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $, whereas observed critical 
Reynolds number\cite{Grizzle96} (Grizzle 1996) for monami is much higher $\approx O(1000)$. These drawbacks of existing theory suggest that flow through vegetation requires 
further investigation for better understanding of phenomenon.
In this paper we extend the idea of Raupach et.al\cite{Raupach96} via incorporating drag due to presence of vegetation. Our calculation of linear stability analysis of flow in presence of grass
shows that flow through vegetation become unstable not only through a mechanism of Kelvin-Helmholtz 
instability associated with strong shear near the canopy top but also through shear instability of flow above grass canopies and shear instability of flow comprising flow above the canopy 
and flow in the shear layer. We further show that inclusion of drag due to grass predicts a threshold condition for waving. We confirm our analysis by comparing prediction from theory 
with field and experimental observations.
%\section{The model}
%Monami is believed to be response to strong oscillations of flow in streamwise velocity associated with coherent vortices.

>>>>

\begin{document}
\title{Waving and swaying}
\author{Ravi Singh}
\affiliation{Brown University, Providence RI 02912 USA}
\author{L. Mahadevan}
\affiliation{Harvard University, Cambridge MA 02138 USA}
\author{Mahesh Bandi\footnote{work performed while visiting Brown University}}
\affiliation{Okinawa Institute of Science and Technology, Okinawa, Japan}
\author{Amala Mahadevan}
\affiliation{Woods Hold Institute of Oceanography, Woods Hole MA USA}
\author{Shreyas Mandre}
\affiliation{Brown University, Providence RI 02912 USA}

@@

\section{Introduction}
Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as 
transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc \cite{Fonseca87,Nepf99}. 
The response of flexible grass canopies to steady current in form of large amplitude coherent oscillation, known as mo-nami \cite{AckermanOkubo93}, plays a central role
in the recruitment of microscopic marine organism such as blue mussel larvae \cite{Grizzle96}. The hydrodynamic mechanism underlying monami is the focus of this paper. 
\newline
Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically. Another major
difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies.
Due to these differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domian is applicable only to marine canopies. 
\newline   
Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to
different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to 
Kelvin-Helmholtz instability. Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
%Sea grasses exhibit large amplitude wave like motion in response to these eddies propagating downstream.
%are exhibited in form of large amplitude oscillations knows as mo-nami.
\newline
While shear layer model successfully appears to predict frequency of mo-nami for number of experimental observations,
several aspect of existing theory remain unsatisfactory. First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self 
consistent theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $, whereas observed critical 
Reynolds number\cite{Grizzle96} (Grizzle 1996) for monami is much higher $\approx O(1000)$. Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability. These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of phenomenon.

In this paper we present a mathematical model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations. Our calculation of linear stability analysis of flow in presence of grass shows that flow through vegetation become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with strong shear near the canopy top but also through an instability of the unvegetated flow coupled with flow deep within the vegetation. We call this new instability the unvegetated layer instability for brevity. We further show that inclusion of drag due to grass predicts a threshold condition for waving. We confirm our analysis by comparing prediction from theory with field and experimental observations.
